---
title: "Tidyverse"
output: html_notebook
---

#5.2 Exercises
1. Examine the built-in dataset co2. Which of the following is true:

A. co2 is tidy data: it has one year for each row.
B. co2 is not tidy: we need at least one column with a character vector.
C. co2 is not tidy: it is a matrix instead of a data frame.
D. co2 is not tidy: to be tidy we would have to wrangle it to have three columns (year, month and value), then each co2 observation would have a row.
```{r}
library(tidyverse)
```
```{r}
data(co2)
head(co2)
```
A> C

2. Examine the built-in dataset ChickWeight. Which of the following is true:

A. ChickWeight is not tidy: each chick has more than one row.

B. ChickWeight is tidy: each observation (a weight) is represented by one row. The chick from which this measurement came from is one the variables.

C. ChickWeight is not a tidy: we are missing the year column.

D. ChickWeight is tidy: it is stored in a data frame.
```{r}
head(ChickWeight)
```
A> B

#. 5.4 Exercise
1. 
```{r}
library(dplyr)
library(dslabs)
data(murders)
murders <- mutate(murders, population_in_millions = population / 10^6)
head(murders)
```
```{r}
murders <- mutate(murders, rate = total / population * 100000)
head(murders)
```

2. If rank(x) gives you the ranks of x from lowest to highest, rank(-x) gives you the ranks from highest to lowest. Use the function mutate to add a column rank containing the rank, from highest to lowest murder rate. Make sure you redefine murders so we can keep using this variable.
```{r}
murders <- mutate(murders, rank = rank(-murders$rate))
head(murders)
```

3. With dplyr, we can use select to show only certain columns. For example, with this code we would only show the states and population sizes: