---
title: "Tidyverse"
output: html_notebook
---

#5.2 Exercises
1. Examine the built-in dataset co2. Which of the following is true:

A. co2 is tidy data: it has one year for each row.
B. co2 is not tidy: we need at least one column with a character vector.
C. co2 is not tidy: it is a matrix instead of a data frame.
D. co2 is not tidy: to be tidy we would have to wrangle it to have three columns (year, month and value), then each co2 observation would have a row.
```{r}
library(tidyverse)
```
```{r}
data(co2)
head(co2)
```
A> C

2. Examine the built-in dataset ChickWeight. Which of the following is true:

A. ChickWeight is not tidy: each chick has more than one row.

B. ChickWeight is tidy: each observation (a weight) is represented by one row. The chick from which this measurement came from is one the variables.

C. ChickWeight is not a tidy: we are missing the year column.

D. ChickWeight is tidy: it is stored in a data frame.
```{r}
head(ChickWeight)
```
A> B

#. 5.4 Exercise
1. 
```{r}
library(dplyr)
library(dslabs)
data(murders)
murders <- mutate(murders, population_in_millions = population / 10^6)
head(murders)
```
```{r}
murders <- mutate(murders, rate = total / population * 100000)
head(murders)
```

2. If rank(x) gives you the ranks of x from lowest to highest, rank(-x) gives you the ranks from highest to lowest. Use the function mutate to add a column rank containing the rank, from highest to lowest murder rate. Make sure you redefine murders so we can keep using this variable.
```{r}
murders <- mutate(murders, rank = rank(-murders$rate))
head(murders)
```

3. With dplyr, we can use select to show only certain columns. For example, with this code we would only show the states and population sizes:
Use select to show the state names and abbreviations in murders. Do not redefine murders, just show the results.
```{r}
select(murders, state, abb) %>% head()
```

4. The dplyr function filter is used to choose specific rows of the data frame to keep. Unlike select which is for columns, filter is for rows. For example, you can show just the New York row like this:
You can use other logical vectors to filter rows.

Use filter to show the top 5 states with the highest murder rates. After we add murder rate and rank, do not change the murders dataset, just show the result. Remember that you can filter based on the rank column.
```{r}
filter(murders, rank <= 5)
```

5. We can remove rows using the != operator. For example, to remove Florida, we would do this:
no_florida <- filter(murders, state != "Florida")
Create a new data frame called no_south that removes states from the South region. How many states are in this category? You can use the function nrow for this.
```{r}
no_south <- filter(murders, region != "South")
head(no_south)
```

6. We can also use %in% to filter with dplyr. You can therefore see the data from New York and Texas like this:
filter(murders, state %in% c("New York", "Texas"))
Create a new data frame called murders_nw with only the states from the Northeast and the West. How many states are in this category?
```{r}
murders_nw <- filter(murders, region %in% c("Northeast", "West"))
tail(murders_nw)
```

7. Suppose you want to live in the Northeast or West and want the murder rate to be less than 1. We want to see the data for the states satisfying these options. Note that you can use logical operators with filter. Here is an example in which we filter to keep only small states in the Northeast region.
Make sure murders has been defined with rate and rank and still has all states. Create a table called my_states that contains rows for states satisfying both the conditions: it is in the Northeast or West and the murder rate is less than 1. Use select to show only the state name, the rate and the rank.
```{r}
filter(murders_nw, rate < 1)
```

OR also it can be like this:
```{r}
filter(murders, region %in% c("Northeast", "West") & rate < 1)
```

#. 5.6 Exercises
1. The pipe %>% can be used to perform operations sequentially without having to define intermediate objects. Start by redefining murder to include rate and rank.
Repeat the previous exercise, but now instead of creating a new object, show the result and only include the state, rate, and rank columns. Use a pipe %>% to do this in just one line.
```{r}
murders %>% select(state, rate, rank) %>% filter(rank <= 5)
```

2. Reset murders to the original table by using data(murders). Use a pipe to create a new data frame called my_states that considers only states in the Northeast or West which have a murder rate lower than 1, and contains only the state, rate and rank columns. The pipe should also have four components separated by three %>%. The code should look something like this:
```{r}
data(murders)
head(murders)
```

```{r}
my_state <- murders %>%
  mutate(rate = total / population * 100000) %>%
  mutate(rank = rank(-rate)) %>%
  filter(rank <= 5) %>%
  select(state, rate, rank)
data.frame(my_state)
```

```{r}
library(dplyr)
library(dslabs)
data(heights)
```


We can obtain these three values with just one line using the quantile function: for example, quantile(x, c(0,0.5,1)) returns the min (0th percentile), median (50th percentile), and max (100th percentile) of the vector x. However, if we attempt to use a function like this that returns two or more values inside summarize:
heights %>% 
  filter(sex == "Female") %>%
  summarize(range = quantile(height, c(0, 0.5, 1)))
we will receive an error: Error: expecting result of length one, got : 2. With the function summarize, we can only call functions that return a single value. In Section 5.12, we will learn how to deal with functions that return more than one value.

since, as most dplyr functions, summarize always returns a data frame.

This might be problematic if we want to use this result with functions that require a numeric value. Here we show a useful trick for accessing values stored in data when using pipes: when a data object is piped that object and its columns can be accessed using the pull function. To understand what we mean take a look at this line of code:

us_murder_rate %>% pull(rate)
The result does not look very different from heights, except we see Groups: sex [2] when we print the object. Although not immediately obvious from its appearance, this is now a special data frame called a grouped data frame and dplyr functions, in particular summarize, will behave differently when acting on this object. Conceptually, you can think of this table as many tables, with the same columns but not necessarily the same number of rows, stacked together in one object. When we summarize the data after grouping, this is what happens:

heights %>% 
  group_by(sex) %>%
  summarize(average = mean(height), standard_deviation = sd(height))
A tibble: 2 x 3
sex    average standard_deviation
<fct>    <dbl>              <dbl>
1 Female    64.9               3.76
2 Male      69.3               3.61

The summarize function applies the summarization to each group separately.  

When examining a dataset, it is often convenient to sort the table by the different columns. We know about the order and sort function, but for ordering entire tables, the dplyr function arrange is useful. For example, here we order the states by population size:

Note that the default behavior is to order in ascending order. In dplyr, the function desc transforms a vector so that it is in descending order. To sort the table in descending order, we can type:

murders %>% 
  arrange(desc(rate)) %>% 

In the code above, we have used the function head to avoid having the page fill up with the entire dataset. If we want to see a larger proportion, we can use the top_n function. This function takes a data frame as it’s first argument, the number of rows to show in the second, and the variable to filter by in the third.
murders %>% top_n(10, rate)


#. 5.9 Exercises
```{r}
library(NHANES)
data(NHANES)
head(NHANES)
```

1. We will provide some basic facts about blood pressure. First let’s select a group to set the standard. We will use 20-29 year old females. AgeDecade is a categorical variable with these ages. Note that the category is coded like " 20-29“, with a space in front! What is the average and standard deviation of systolic blood pressure as saved in the BPSysAve variable? Save it to a variable called ref.
```{r}
ref <- NHANES %>%
  filter (AgeDecade == " 20-29") %>%
  group_by(Gender) %>%
  summarize(average = mean(BPSysAve, na.rm = TRUE), standard_deviation = sd(BPSysAve, na.rm = TRUE))
ref
```

2. Using a pipe, assign the average to a numeric variable ref_avg. Hint: Use the code similar to above and then pull.
```{r}
ref_avg. <- ref %>% pull(average)
class(ref_avg.)
```

3. Report the min and max values for the same group.
```{r}
NHANES %>%
  filter (AgeDecade == " 20-29") %>%
  group_by(Gender) %>%
  summarize(minimum = min(BPSysAve, na.rm = TRUE), maximum = max(BPSysAve, na.rm = TRUE))
```

4. Compute the average and standard deviation for females, but for each age group separately rather than a selected decade as in question 1. Note that the age groups are defined by AgeDecade. Hint: rather than filtering by age and gender, filter by Gender and then use group_by.
```{r}
NHANES %>%
  filter (Gender == "female") %>%
  group_by(AgeDecade) %>%
  summarize(average = mean(BPSysAve, na.rm = TRUE), standard_deviation = sd(BPSysAve, na.rm = TRUE))
```

5. Repeat exercise 4 for males.
```{r}
NHANES %>%
  filter (Gender == "male") %>%
  group_by(AgeDecade) %>%
  summarize(average = mean(BPSysAve, na.rm = TRUE), standard_deviation = sd(BPSysAve, na.rm = TRUE))
```

6. We can actually combine both summaries for exercises 4 and 5 into one line of code. This is because group_by permits us to group by more than one variable. Obtain one big summary table using group_by(AgeDecade, Gender).
```{r}
NHANES %>%
  group_by(Gender, AgeDecade) %>%
  summarize(average = mean(BPSysAve, na.rm = TRUE), standard_deviation = sd(BPSysAve, na.rm = TRUE))
```

7. For males between the ages of 40-49, compare systolic blood pressure across race as reported in the Race1 variable. Order the resulting table from lowest to highest average systolic blood pressure.
```{r}
NHANES %>%
  filter(Gender == "male" & AgeDecade == " 40-49") %>%
  group_by(Race1) %>%
  arrange(BPSysAve) 
```

#. 5.10 Tibbles
The tbl, pronounced tibble, is a special kind of data frame. The functions group_by and summarize always return this type of data frame. The group_by function returns a special kind of tbl, the grouped_df. We will say more about these later. For consistency, the __dplyr__manipulation verbs (select, filter, mutate, and arrange) preserve the class of the input: if they receive a regular data frame they return a regular data frame, while if they receive a tibble they return a tibble. But tibbles are the preferred format in the tidyverse and as a result tidyverse functions that produce a data frame from scratch return a tibble. For example, in Chapter 6 we will see that tidyverse functions used to import data create tibbles.

Tibbles are very similar to data frames. In fact, you can think of them as a modern version of data frames. Nonetheless there are three important differences which we describe in the next.

##. 5.10.1 Tibbles display better
The print method for tibbles is more readable than that of a data frame. To see this, compare the outputs of typing murders and the output of murders if we convert it to a tibble. We can do this using as_tibble(murders). If using RStudio, output for a tibble adjusts to your window size. To see this, change the width of your R console and notice how more/less columns are shown.
```{r}
library(tidyverse)
data(murders)
as_tibble(murders)
```

##. 5.10.3 Tibbles can have complex entries
While data frame columns need to be vectors of numbers, strings or logical values, tibbles can have more complex objects, such as lists or functions. Also, we can create tibbles with functions:
```{r}
tibble(id = c(1, 2, 3), func = c(mean, median, sd))

```

##. 5.10.5 Create a tibble using data_frame instead of data.frame
It is sometimes useful for us to create our own data frames. To create a data frame in the tibble format, you can do this by using the data_frame function.
```{r}
grades <- data_frame(names = c("John", "Juan", "Jean", "Yao"), exam_1 = c(95, 80, 90, 85), exam_2 = c(90, 85, 85, 90))
grades
```

#. 5.11 The dot operator
One of the advantages of using the pipe %>% is that we do not have to keep naming new objects as we manipulate the data frame. 

We can do this because each of these functions takes a data frame as the first argument. But what if we want to access a component of the data frame. For example, what if the pull function was not available and we wanted to access tab_2$rate? What data frame name would we use? The answer is the dot operator.

```{r}
rates <- filter(murders, region == "Northeast") %>%
  mutate(rate = total / population * 10^5) %>%
  .$rate
max(rates)
```

#. 5.12 do
The tidyverse functions know how to interpret grouped tibbles. Furthermore, to facilitate stringing commands through the pipe %>%, tidyverse functions consistently return data frames, since this assures that the output of a function is accepted as the input of another. But most R functions do not recognize grouped tibbles nor do they return data frames. The quantile function is an example we described in Section 5.7.1. The do functions serves as a bridge between R functions such as quantile and the tidyverse. The do function understands grouped tibbles and always returns a data frame.

In Section 5.7.1, we noted that if we attempt to use quantile to obtain the min, median and max in one call, we will receive an error: Error: expecting result of length one, got : 2.

We can use the do function fix this.

First we have to write a function that fits into the tidyverse approach: that is, it receives a data frame and returns a data frame.
```{r}
my_summary <- function(dat){
  x <- quantile(dat$height, c(0, 0.5, 1))
  data_frame(min = x[1], median = x[2], max = x[3])
}

heights %>%
  group_by(sex) %>%
  my_summary
```

But this is not what we want. We want a summary for each sex and the code returned just one summary. This is because my_summary is not part of the tidyverse and does not know how to handled grouped tibbles. do makes this connection:

```{r}
heights %>%
  group_by(sex) %>%
  do(my_summary(.))
```
Note that here we need to use the dot operator. The tibble created by group_by is piped to do. Within the call to do, the name of this tibble is . and we want to send it to my_summary. If you do not use the dot, then my_summary has __no argument and returns an error telling us that argument "dat" is missing.
If you do not use the parenthesis, then the function is not executed and instead do tries to return the function. This gives an error because do must always return a data frame. 

#. 5.13 The purrr package
```{r}
compute_s_n <- function(n){
  x <- 1:n
  sum(x)
}
n <- 1:25
s_n <- sapply(n, compute_s_n)
a_s_n <- as.numeric(s_n)
class(a_s_n)
```

This type of operation, applying the same function or procedure to elements of an object, is quite common in data analysis. The purrr package includes functions similar to sapply but that better interact with other tidyverse functions. The main advantage is that we can better control the output type of functions. In contrast, sapply can return several different object types; for example, we might expect a numeric result from a line of code, but sapply might convert our result to character under some circumstances. purrr functions will never do this: they will return objects of a specified type or return an error if this is not possible.
The first purrr function we will learn is map, which works very similar to sapply but always, without exception, returns a list. If we want a numeric vector, we can instead use map_dbl which always returns a vector of numeric values. 
A particularly useful purrr function for interacting with the rest of the tidyverse is map_df, which always returns a tibble data frame. 

*sapply는 integer/character 등으로 만드는데, map (purrr의 일종)은 list로 나옴. 


#. 5.14 Tidyverse conditionals
A typical data analysis will often involve one or more conditional operation. In Section 4.1 we described the ifelse function, which we will use extensively in this book. In this section we present two dplyr functions that provide further functionality for performing conditional operations.

##. 5.14.1 case_when
The case_when function is useful for vectorizing conditional statements. It is similar to ifelse but can output any number of values, as opposed to just TRUE or FALSE. Here is an example splitting numbers into negative, positives and 0:
```{r}
x <- c(-2, -1, 0, 1, 2)
case_when(x < 0 ~ "Negative", x > 0 ~ "Positive", TRUE ~ "Zero")
```

```{r}
data(murders)
murders %>%
  mutate(group = case_when(abb %in% c("ME", "NH", "VT", "MA", "RI", "CT") ~ "New England", abb %in% c("WA", "OR", "CA") ~ "West Coast", region == "South" ~ "South", TRUE ~ "others")) %>%
  group_by(group) %>%
  summarize(rate = sum(total) / sum(population) * 10^5)
```

##. 5.14.2 between
A common operation in data analysis is to determine if a value falls inside an interval. We can check this using conditionals. For example to check if the elements of a vector x are between a and b we can type

x >= a & x <= b
However, this can become cumbersome, especially within the tidyverse approach. The between function performs the same operation.

between(x, a, b)

#. 5.15 Exercises
1. Load the murders dataset. Which of the following is true?

A. murders is in tidy format and is stored in a tibble.
B. murders is in tidy format and is stored in a data frame. <- answer
C. murders is not in tidy format and is stored in a tibble.
D. murders is not in tidy format and is stored in a data frame.
```{r}
data(murders)
head(murders)
```

2. Use as_tibble to covert the murders data table into a tibble and save it in an object called murders_tibble.
```{r}
murders_tibble <- as_tibble(murders)
```

3. Use the group_by function to convert murders into a tibble that is grouped by region.
```{r}
murders %>%
  as_tibble(murders) %>%
  group_by(region)
```

4. Write tidyverse code that is equivalent to this code: exp(mean(log(murders$population))). Write it using the pipe so that each function is called without arguments. Use the dot operator to access the population. Hint: The code should start with murders %>%.
```{r}
murders %>%
  mutate(logarithm = log(population)) %>%
  summarize(avg = exp(mean(logarithm))) %>%
  pull(avg)

murders %>%
  mutate(logarithm = log(population)) %>%
  .$logarithm
exp(mean(logarithm))
```
dot function 사용 방법 모르겠음...

5. Use the map_df to create a data frame with three columns named n, s_n, and s_n_2. The first column should contain the numbers 1 through 100. The second and third columns should each contain the sum of 1 through n with n the row number.
```{r}
compute_s_n <- function(n){
  x <- 1:n
  data_frame(n = n, s_n = sum(x), s_n_2 = sum(x))
}
n <- 1:100
s_n <- map_df(n, compute_s_n)
head(s_n)

```

